Can we use masked priming in an online setting?

### 2021-05-19 MaskedPriming_complete_mspg.Rmd
Update MaskedPriming_complete_mspg.Rmd

abstract: |
  Masked priming is one of the most important paradigms in the study of visual word recognition, but it is usually thought to require a laboratory setup with a known monitor and keyboard. To test if this technique can be safely used in an online setting, we conducted two online masked priming lexical decision task experiments using PsychoPy/PsychoJS [@peirce2019]. Importantly, we also tested the role of prime exposure duration (33.3 vs. 50 ms in Experiment 1 and 16.7 vs. 33.3 ms in Experiment 2), thus allowing us to examine both across conditions and within-conditions effects. We found that our online data are indeed very similar to the masked priming data reported by @gomezDiffusionModelAccount2013. Additionally, we found a clear effect of prime duration, with the priming effect (measured in terms of response time and accuracy) being stronger at 50 ms than 33.3 ms and no priming effect at 16.7 ms prime duration. From these results, we can conclude that modern online browser-based experimental psychophysics packages (e.g., PsychoPy) can present stimuli and collect responses on standard end user devices with enough precision. These findings provide us with confidence that masked priming can be used online, thus allowing us not only to run less time-consuming experiments, but also to reach populations that are difficult to test in a laboratory.

@@

Cognitive psychologists have been much slower that their behavioral economics coleagues in taking up online paradigms [see @brysbaert2016impact; @cai2017accent;@eerland2013influence; @rodd2016impact, for some exceptions], often due to concerns about the validity of the results. Such concerns are not limited to cognitive studies [see @austSeriousnessChecksAre2013], but they are exacerbated by the reliance on precise presentation times in cognitive psychology. Furthermore, these concerns are even more central in the masked priming technique, where it is critical for the onset of the mask, the prime, and the target to occur at the nominal times. For instance, presenting the masked prime for longer than intended (e.g., 83 ms or longer instead of the nominal 50 ms) could counteract the effect of the mask, making the prime consciously visible to the participant and possibly altering the processes of interest [see @Zimmerman_2012].

@@

Descriptive statistics for response times and accuracy for both words and nonwords in the experimental conditions are reported in Table~\@ref(tab:exp1-descriptive-rt-table) (although note that we only analyzed the word trials).

@@

Table~\@ref(tab:exp1-blmm-table) shows the mean, standard error, lower and upper bounds of the 95% CrI of the estimate of each fixed effect in the RT model, as well as the $\hat{R}$ for each estimate, which indicate that the model was fitted successfully as they are all close to 1.

@@

Table~\@ref(tab:exp1-acc-blmm-table) shows the mean, standard error, lower and upper bounds of the 95% CrI of the estimate of each fixed effect in the accuracy model, as well as the $\hat{R}$ for each estimate, which indicate that the model was fitted successfully as they are all close to 1.

@@

Descriptive statistics for RTs and accuracy in all the experimental conditions are reported in Table~\@ref(tab:exp2-descriptive-rt-table).

@@

apa_table(
    blmm_exp2_rt_fixed_for_print,
    placement = "H",
    align = c("c", "c", "c", "c", "c"),
    caption = "(ref:exp2-blmm-rt-table-caption)",
    note = "$\\beta$ is the scale parameter (the inverse of the rate parameter $\\lambda$) of the ex-Gaussian distribution.",
    small = TRUE,
    digits = 3,
    escape = FALSE
    )

@@

Furthermore, our results extend previous work showing the validity of on-line studies measuring response times [e.g., @brysbaert2016impact; @cai2017accent; @eerland2013influence; @rodd2016impact, to cite a few instances] to the masked priming technique. Thus, the present study opens the door to a wider use of online experiments in cognitive research, especially in reaction-time sensitive fields like word recognition. Not only does this enable researchers to continue collecting data in times of social distancing, but it also makes it possible to collect data from a larger population than previously possible. For instance, people from different countries and/or cultures, from different age groups, bilinguals/multilinguals, even from those who do not have computers at home, but they do have smartphones. Using a Javascript-based experiment software, anyone with a smartphone can be an experiment participant. Additionally, deploying masked priming experiments online also means that data can be collected very quickly and efficiently, allowing research to progress more rapidly.

>>>>

abstract: |
  Masked priming is one of the most important paradigms in the study of visual word recognition, but it is usually thought to require a laboratory setup with a known monitor and keyboard. To test if this technique can be safely used in an online setting, we conducted two online masked priming lexical decision task experiments using PsychoPy/PsychoJS [@peirce2019]. Importantly, we also tested the role of prime exposure duration (33.3 vs. 50 ms in Experiment 1 and 16.7 vs. 33.3 ms in Experiment 2), thus allowing us to examine both across conditions and within-conditions effects. We found that our online data are indeed very similar to the masked priming data previously reported in the masked priming literature. Additionally, we found a clear effect of prime duration, with the priming effect (measured in terms of response time and accuracy) being stronger at 50 ms than 33.3 ms and no priming effect at 16.7 ms prime duration. From these results, we can conclude that modern online browser-based experimental psychophysics packages (e.g., PsychoPy) can present stimuli and collect responses on standard end user devices with enough precision. These findings provide us with confidence that masked priming can be used online, thus allowing us not only to run less time-consuming experiments, but also to reach populations that are difficult to test in a laboratory.

@@

Cognitive psychologists have been much slower that their behavioral economics colleagues in taking up online paradigms [see @brysbaert2016impact; @cai2017accent; @Dufau_2011; @eerland2013influence; @rodd2016impact, for some exceptions], often due to concerns about the validity of the results. Such concerns are not limited to cognitive studies [see @austSeriousnessChecksAre2013], but they are exacerbated by the reliance on precise presentation times in cognitive psychology. Furthermore, these concerns are even more central in the masked priming technique, where it is critical for the onset of the mask, the prime, and the target to occur at the nominal times. For instance, presenting the masked prime for longer than intended (e.g., 83 ms or longer instead of the nominal 50 ms) could counteract the effect of the mask, making the prime consciously visible to the participant and possibly altering the processes of interest [see @Zimmerman_2012].

@@

Descriptive statistics for response times and accuracy for both words and nonwords in the experimental conditions are reported in Table \@ref(tab:exp1-descriptive-rt-table) (although note that we only analyzed the word trials).

@@

Table \@ref(tab:exp1-blmm-table) shows the mean, standard error, lower and upper bounds of the 95% CrI of the estimate of each fixed effect in the RT model, as well as the $\hat{R}$ for each estimate, which indicate that the model was fitted successfully as they are all close to 1.

@@

Table \@ref(tab:exp1-acc-blmm-table) shows the mean, standard error, lower and upper bounds of the 95% CrI of the estimate of each fixed effect in the accuracy model, as well as the $\hat{R}$ for each estimate, which indicate that the model was fitted successfully as they are all close to 1.

@@

Descriptive statistics for RTs and accuracy in all the experimental conditions are reported in Table \@ref(tab:exp2-descriptive-rt-table).

@@

apa_table(
    blmm_exp2_rt_fixed_for_print,
    placement = "H",
    align = c("c", "c", "c", "c", "c"),
    caption = "(ref:exp2-blmm-table-caption)",
    note = "$\\beta$ is the scale parameter (the inverse of the rate parameter $\\lambda$) of the ex-Gaussian distribution.",
    small = TRUE,
    digits = 3,
    escape = FALSE
    )

@@

Furthermore, our results extend previous work showing the validity of on-line studies measuring response times [e.g., @brysbaert2016impact; @cai2017accent; @Dufau_2011; @eerland2013influence; @rodd2016impact, to cite a few instances] to the masked priming technique. Thus, the present study opens the door to a wider use of online experiments in cognitive research, especially in reaction-time sensitive fields like word recognition. Not only does this enable researchers to continue collecting data in times of social distancing, but it also makes it possible to collect data from a larger population than previously possible. For instance, people from different countries and/or cultures, from different age groups, bilinguals/multilinguals, even from those who do not have computers at home, but they do have smartphones. Using a Javascript-based experiment software, anyone with a smartphone can be an experiment participant. Additionally, deploying masked priming experiments online also means that data can be collected very quickly and efficiently, allowing research to progress more rapidly.
